### 2023-09-05 proofs/tex/equiv-lemmas.tex
Rearrange proofs

\begin{lemma}[Algorithmization of declarative equivalence]
  \label{lemma:decl-equiv-algorithmization}
  Declarative equivalence is the equality of normal forms. 
  \begin{itemize}
    \item[$+$] $[[iP ≈ iQ]] \iff [[nf(iP) = nf(iQ)]]$,
    \item[$-$] $[[iN ≈ iM]] \iff [[nf(iN) = nf(iM)]]$.
  \end{itemize}
\end{lemma}
\begin{proof} \hfill
  \begin{itemize}
    \item[$+$] Let us prove both directions separately.
    \begin{itemize}
      \item[$\Rightarrow$] 
        exactly by \cref{lemma:normalization-completeness},
      \item[$\Leftarrow$] 
        from \cref{lemma:normalization-soundness}, we know
        $[[iP ≈ nf(iP)]] = [[nf(iQ) ≈ iQ]]$, then by transitivity (\cref{lemma:decl-equiv-transitivity}),
        $[[iP ≈ iQ]]$.
    \end{itemize}
    \item[$-$] The proof is exactly the same.
  \end{itemize}
\end{proof}

@@

\begin{lemma}[Algorithmic type well-formedness is invariant under equivalence]
  \label{lemma:wf-equiv-algo}
  Mutual subtyping implies declarative equivalence.
  \begin{itemize}
  \item[$+$] if $[[uP ≈ uQ]]$ then $[[Γ; Ξ ⊢ uP]] \iff [[Γ; Ξ ⊢ uQ]]$,
  \item[$-$] if $[[uN ≈ uM]]$ then $[[Γ; Ξ ⊢ uN]] \iff [[Γ; Ξ ⊢ uM]]$
  \end{itemize}
\end{lemma}
\begin{proof}
  The proof coincides with the proof of \cref{lemma:wf-equiv},
  and adds two cases for equating two positive or two negative metavariables,
  which must be equal by inversion, and thus, 
  $[[Γ; Ξ ⊢ α̂±]] \iff [[Γ; Ξ  ⊢ α̂±]]$ holds trivially.
\end{proof}

@@

@@

\begin{corollary}[Normalization preserves well-formedness of algorithmic types]
  \label{corollary:wf-nf-algo}
  \begin{itemize}
    \item[$+$] $[[Γ; Ξ ⊢ uP]] \iff [[Γ; Ξ ⊢ nf(uP)]]$,
    \item[$-$] $[[Γ; Ξ ⊢ uN]] \iff [[Γ; Ξ ⊢ nf(uN)]]$
  \end{itemize}
\end{corollary}
\begin{proof}
  Immediately from \cref{lemma:wf-equiv-algo,lemma:normalization-soundness-alg}.
\end{proof}

\begin{corollary}[Normalization preserves well-formedness of substitution]
  \label{corollary:wf-s-nf}
  \hfill \\
   $[[Γ2 ⊢ σ : Γ1]] \iff [[Γ2 ⊢ nf(σ) : Γ1]]$
\end{corollary}
\begin{proof}
  Let us prove the forward direction.
  Suppose that $[[α± ∊ Γ1]]$.  Let us show that $[[Γ2 ⊢ [nf(σ)]α±]]$.
  By the definition of substitution normalization,
  $[[ [nf(σ)]α± = nf([σ]α±) ]]$. Then by \cref{corollary:wf-nf-algo},
  to show $[[Γ2 ⊢ nf([σ]α±)]]$, it suffices to show $[[Γ2 ⊢ [σ]α±]]$,
  which holds by the assumption $[[Γ2 ⊢ σ : Γ1]]$.

\begin{corollary}[Normalization preserves the signature of the algorithmic substitution]
  \label{corollary:norm-subst-sig-algo}
  \hfill \\
  $[[Θ ⊢ uσ : Ξ]] \iff [[Θ ⊢ nf(uσ) : Ξ]]$,
  $[[Γ ⊢ uσ : Ξ]] \iff [[Γ ⊢ nf(uσ) : Ξ]]$.
\end{corollary}
\begin{proof}
  The proof is analogous to \cref{corollary:wf-s-nf}.
\end{proof}

@@

\begin{lemma}[Algorithmization of subtyping-induced equivalence]
  \label{lemma:subt-equiv-algorithmization}
  Mutual subtyping is equality of normal forms.
 Assuming all the types below are well-formed in $[[Γ]]$:
  \begin{itemize}
    \item [$+$] $[[Γ ⊢ iP ≈ iQ]] \iff [[nf(iP) = nf(iQ)]]$,
    \item [$-$] $[[Γ ⊢ iN ≈ iM]] \iff [[nf(iN) = nf(iM)]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
  Let us prove the positive case, the negative case is symmetric.
  We prove both directions of $\iff$ separately:
  \begin{itemize}
    \item [$\Rightarrow$] exactly \cref{corollary:nf-complete-wrt-subt-equiv};
    \item [$\Leftarrow$] by \cref{lemma:decl-equiv-algorithmization,lemma:equiv-soundness}.
  \end{itemize}
\end{proof}

\begin{corollary}[Algorithmic substitution equivalence becomes equality after normalization]
  \label{corollary:subst-subt-equiv-algorithmization}
  Suppose that $[[Θ ⊢ uσ1 : Ξ']]$ and $[[Θ ⊢ uσ2 : Ξ']]$ are algorithmic substitutions
  and $[[Ξ ⊆ Ξ']]$. 
  Then $[[Γ ⊢ uσ1 ≈ uσ2 : Ξ]] \iff [[nf(uσ1)|Ξ = nf(uσ2)|Ξ]]$.
\end{corollary}

>>>>

@@

@@

\begin{corollary}[Normalization preserves well-formedness of substitution]
  \label{corollary:wf-s-nf}
  \hfill \\
   $[[Γ2 ⊢ σ : Γ1]] \iff [[Γ2 ⊢ nf(σ) : Γ1]]$
\end{corollary}
\begin{proof}
  Let us prove the forward direction.
  Suppose that $[[α± ∊ Γ1]]$.  Let us show that $[[Γ2 ⊢ [nf(σ)]α±]]$.
  By the definition of substitution normalization,
  $[[ [nf(σ)]α± = nf([σ]α±) ]]$. Then by \cref{corollary:wf-nf},
  to show $[[Γ2 ⊢ nf([σ]α±)]]$, it suffices to show $[[Γ2 ⊢ [σ]α±]]$,
  which holds by the assumption $[[Γ2 ⊢ σ : Γ1]]$.

@@

@@

\begin{lemma}[Algorithmization of subtyping-induced equivalence]
  \label{lemma:subt-equiv-algorithmization}
  Mutual subtyping is the equality of normal forms.
 Assuming all the types below are well-formed in $[[Γ]]$:
  \begin{itemize}
    \item [$+$] $[[Γ ⊢ iP ≈ iQ]] \iff [[nf(iP) = nf(iQ)]]$,
    \item [$-$] $[[Γ ⊢ iN ≈ iM]] \iff [[nf(iN) = nf(iM)]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
  Let us prove the positive case, the negative case is symmetric.
  We prove both directions of $\iff$ separately:
  \begin{itemize}
    \item [$\Rightarrow$] exactly \cref{corollary:nf-complete-wrt-subt-equiv};
    \item [$\Leftarrow$] by \cref{lemma:decl-equiv-algorithmization,lemma:equiv-soundness}.
  \end{itemize}
\end{proof}

### 2023-09-05 proofs/tex/norm-alg-lemmas.tex
Rearrange proofs

\begin{lemma}
  \label{lemma:equiv-fv}
  Set of free variables is invariant under equivalence.
  \begin{itemize}
  \item[$-$] If $[[iN ≈ iM]]$ then $[[fv iN]] \equiv [[fv iM]]$ (as sets)
  \item[$+$] If $[[iP ≈ iQ]]$ then $[[fv iP]] \equiv [[fv iQ]]$ (as sets)
  \end{itemize}
\end{lemma}
\begin{proof}
  Straightforward mutual induction on $[[iN ≈ iM]]$ and $[[iP ≈ iQ]]$
\end{proof}

\begin{lemma}
  \label{lemma:fv-nf}
  Free variables are not changed by the normalization
  \begin{itemize}
  \item[$-$] $[[fv iN]] \equiv [[fv nf(iN)]]$
  \item[$+$] $[[fv iP]] \equiv [[fv nf(iP)]]$
  \end{itemize}
\end{lemma}
\begin{proof}
  By straightforward induction on $[[iN]]$ and mutually on $[[iP]]$.
\end{proof}

\begin{lemma}
  \label{lemma:uv-nf}
  Algorithmic variables are not changed by the normalization
  \begin{itemize}
  \item[$-$] $[[uv uN]] \equiv [[uv nf(uN)]]$
  \item[$+$] $[[uv uP]] \equiv [[uv nf(uP)]]$
  \end{itemize}
\end{lemma}
\begin{proof}
  By straightforward induction on $[[uN]]$ and mutually on $[[uP]]$.
\end{proof}

\begin{lemma}[Soundness of normalization]
  \label{lemma:normalization-soundness}
  \hfill
  \begin{itemize}
    \item[$-$] $[[iN ≈ nf(iN)]]$
    \item[$+$] $[[iP ≈ nf(iP)]]$
  \end{itemize}
\end{lemma}
\begin{proof}
  Mutual induction on $[[nf(iN) = iM]]$ and $[[nf(iP) = iQ]]$.
  Let us consider how this judgment is formed:
  \begin{caseof}
    \item{\nameref{\ottdruleNrmNVarLabel} and \nameref{\ottdruleNrmPVarLabel}}\\ By
      the corresponding equivalence rules.
    \item{\nameref{\ottdruleNrmShiftULabel}, \nameref{\ottdruleNrmShiftDLabel},
        and \nameref{\ottdruleNrmArrowLabel}}\\
      By the induction hypothesis and the corresponding congruent equivalence rules.
    \item{\nameref{\ottdruleNrmForallLabel}}, i.e. $[[nf(∀pas.uN) = ∀pas'.uN']]$ \label{case:norm-soundness:forall}\\
      From the induction hypothesis, we
      know that $[[iN ≈ iN']]$. In particular, by \cref{lemma:equiv-fv}, $[[fv
        iN]] \equiv [[fv iN']]$.
      Then by \cref{lemma:ord-soundness}, $[[{pas'}]]
      \equiv [[{pas} ∩ fv iN']] \equiv [[{pas} ∩ fv iN]]$, and thus,
      $[[{pas'} ∩ fv iN']] \equiv [[{pas} ∩ fv iN]]$.

      To prove $[[∀pas.iN ≈ ∀pas'.iN']]$, it suffices to provide a bijection 
      $\mu : [[{pas'} ∩ fv iN']] \leftrightarrow [[{pas} ∩ fv iN]]$ such that
      $[[iN ≈ [mu]iN']]$. Since these sets are equal, we take $\mu = id$.
    \item{\nameref{\ottdruleNrmExistsLabel}} Same as for \cref{case:norm-soundness:forall}.
  \end{caseof}
\end{proof}

@@

\begin{corollary}[Normalization preserves ordering]
  \label{corollary:normalization-ord}
  For any $[[varset]]$,
  \begin{itemize}
  \item[$-$] $[[ord varset in nf(uN)]] = [[ord varset in uM]]$
  \item[$+$] $[[ord varset in nf(uP)]] = [[ord varset in uQ]]$
  \end{itemize}
\end{corollary}
\begin{proof}
  Immediately from \cref{lemma:ord-completeness,lemma:normalization-soundness}.
\end{proof}

\begin{lemma}[Distributivity of normalization over substitution]
  \label{lemma:norm-subst-distr} Normalization of a term distributes over substitution.
  Suppose that $[[σ]]$ is a substitution, $[[iN]]$ and $[[iP]]$ are types. Then
    \begin{itemize}
      \item[$-$] $[[nf([σ]iN)]] = [[ [nf(σ)] nf(iN) ]]$
      \item[$+$] $[[nf([σ]iP)]] = [[ [nf(σ)] nf(iP) ]]$
  \end{itemize}
  where $[[nf(σ)]]$ means pointwise normalization: $[[ [nf(σ)] α⁻]] = [[nf([σ]α⁻)]]$.
\end{lemma}
\begin{proof}
  Mutual induction on $[[iN]]$ and $[[iP]]$.
  \begin{caseof}
    \item $[[iN]]$ = $[[na]]$ \\
      \label{case:norm-subst-distr-var}
      $[[nf([σ]iN)]] = [[ nf([σ]na) ]] = [[ [nf(σ)]na ]] $.

      $[[ [nf(σ)] nf(iN) ]] = [[ [nf(σ)] nf(na) ]] = [[ [nf(σ)] na ]] $.
    \item $[[iP]]$ = $[[pa]]$ \\
      Similar to \cref{case:norm-subst-distr-var}.
   \item If the type is formed by $[[→]]$, $[[↑]]$, or $[[↓]]$, 
     the required equality follows from the congruence of the normalization and
     substitution, and the induction hypothesis.
     For example, if $[[iN]] = [[iP → iM]]$ then \\
     $\begin{aligned}[t]
        [[nf([σ] iN)]] &= [[ nf([σ] (iP → iM)) ]] \\
                        &= [[ nf([σ]iP → [σ]iM) ]]
                        && \text{By the congruence of substitution} \\
                        &= [[ nf([σ]iP) → nf([σ]iM) ]]
                        && \text{By the congruence of normalization, i.e. \ruleref{\ottdruleNrmArrowLabel}} \\
                        &= [[ [nf(σ)]nf(iP) → [nf(σ)]nf(iM) ]]
                        && \text{By the induction hypothesis} \\
                        &= [[ [nf(σ)](nf(iP) → nf(iM)) ]]
                        && \text{By the congruence of substitution} \\
                        &= [[ [nf(σ)]nf(iP → iM) ]]
                        && \text{By the congruence of normalization} \\
                        &= [[ [nf(σ)]nf(iN) ]]
      \end{aligned}$ \\
    \item $[[iN]] = [[∀ pas.iM]]$ \label{case:norm-subst-commute} \\
      $\begin{aligned}[t]
          [[ [nf(σ)] nf(iN) ]] &= [[ [nf(σ)] nf(∀pas.iM)]] \\
                            &= [[ [nf(σ)] ∀pas'.nf(iM) ]]
                            && \text {Where $[[pas']] = [[ ord {pas} in nf(iM)]]
                               = [[ord {pas} in iM]]$
                               (the latter is by
                               \cref{corollary:normalization-ord})}\\
        \end{aligned}$ \\

      $\begin{aligned}[t]
         [[ nf([σ]iN) ]] &= [[ nf([σ] ∀pas.iM)]] \\
                          &= [[ nf(∀pas.[σ]iM) ]]
                          && \text{Assuming $[[{pas} ∩ Γ1]] = \emptyset$
                             and $[[{pas} ∩ Γ2]] = \emptyset$}\\
                          &= [[ ∀pbs.nf([σ]iM) ]]
                          && \text {Where $[[pbs]] = [[ord {pas} in nf([σ]iM)]]
                             = [[ord {pas} in [σ]iM]]$ (the latter is by
                             \cref{corollary:normalization-ord})}\\
                          &= [[ ∀pas'.nf([σ]iM) ]]
                          && \text{By \cref{lemma:ord-sigma}, $[[pbs]] = [[pas']]$
                             since $[[{pas}]]$ is disjoint with $[[Γ1]]$ and
                             $[[Γ2]]$}\\
                          &= [[ ∀pas'.[nf(σ)]nf(iM) ]]
                          && \text {By the induction hypothesis}\\
         \end{aligned}$ \\

     To show alpha-equivalence of 
     $[[ [nf(σ)] ∀pas'.nf(iM) ]]$ and $[[ ∀pas'.[nf(σ)]nf(iM) ]]$,
     we can assume that $[[{pas'} ∩ Γ1]] = \emptyset$, and $[[{pas'} ∩ Γ2]]
     = \emptyset$.

   \item $[[iP]] = [[∃ nas.iQ]]$ \\
     Same as for \cref{case:norm-subst-commute}.
  \end{caseof}
\end{proof}

\begin{corollary}[Commutativity of normalization and renaming]
  \label{lemma:norm-subst-commute} Normalization of a term commutes with renaming.
  Suppose that $\mu$ is a bijection between two sets of variables
  $\mu : A \leftrightarrow B$. Then
  \begin{itemize}
    \item[$-$] $[[nf([mu]iN)]] = [[ [mu] nf(iN) ]]$
    \item[$+$] $[[nf([mu]iP)]] = [[ [mu] nf(iP) ]]$
  \end{itemize}
\end{corollary}
\begin{proof}
  Immediately from \cref{lemma:norm-subst-distr},
  after noticing that $[[nf(mu)]] = [[mu]]$.
\end{proof}

\begin{lemma}[Completeness of quantified normalization]
  \label{lemma:normalization-completeness}
  Normalization returns the same representative for equivalent types.

  \begin{itemize}
  \item[$-$] If $[[iN ≈ iM]]$ then $[[nf(iN)]] = [[nf(iM)]]$
  \item[$+$] If $[[iP ≈ iQ]]$ then $[[nf(iP)]] = [[nf(iQ)]]$
  \end{itemize}
  (Here equality means alpha-equivalence)
\end{lemma}

\begin{proof}
  Mutual induction on $[[iN ≈ iM]]$ and $[[iP ≈ iQ]]$.
  \begin{caseof}
  \item {\nameref{\ottdruleEOneForallLabel}} \label{case:ord-completeness:forall} \\

    From the definition of the normalization,
    \begin{itemize}
      \item $[[nf(∀pas.iN)]] = [[∀pas'.nf(iN)]]$ where $[[pas']]$ is $[[ord {pas} in nf(iN)]]$
      \item $[[nf(∀pbs.iM)]] = [[∀pbs'.nf(iM)]]$ where $[[pbs']]$ is $[[ord {pbs} in nf(iM)]]$
    \end{itemize}
    Let us take $[[mu : ({pbs} ∩ fv iM) ↔ ({pas} ∩ fv iN)]]$ from the
    inversion of the equivalence judgment. Notice that from
    \cref{lemma:fv-nf,lemma:ord-soundness}, the domain and the codomain of $\mu$ can be written
    as $[[mu : {pbs'} ↔ {pas'}]]$.

    To show the alpha-equivalence of $[[∀pas'.nf(iN)]]$ and $[[∀pbs'.nf(iM)]]$,
    it suffices to prove that
    \begin{enumerate*}
    \item[(i)] $[[ [mu] nf(iM) ]] = [[nf(iN)]]$ and \newline
    \item[(ii)] $[[ [mu]pbs' ]] = [[pas']]$
    \end{enumerate*}.

    \begin{enumerate}
    \item[(i)] $[[ [mu] nf(iM) ]] = [[nf([mu]iM)]] = [[nf(iN)]]$.
      The first equality holds by \cref{lemma:norm-subst-commute}, the second---by the induction hypothesis.

    \item[(ii)] $\begin{aligned}[t] [[ [mu]pbs' ]] &= [[ [mu] ord {pbs} in nf(iM) ]]
                                                  && \text{by the definition of $[[pbs']]$ } \\
                                                  &= [[ [mu] ord ({pbs} ∩ fv iM) in nf(iM) ]]
                                                  && \text{from \cref{lemma:fv-nf,corollary:ord-weakening} } \\
                                                  &= [[ ord [mu] ({pbs} ∩ fv iM) in [mu] nf(iM) ]]
                                                  && \text{by \cref{lemma:distr-mu-ord}, because
                                                     $[[{pas} ∩ fv iN]] \cap [[fv nf(iM)]] \subseteq [[{pas} ∩ fv iM ]]
                                                     = \emptyset$}\\
                                                  &
                                                  && \text{and $[[{pas} ∩ fv iN]] \cap [[({pbs} ∩ fv iM)]] \subseteq
                                                     [[{pas} ∩ fv iM]] = \emptyset$} \\
                                                  &= [[ ord [mu] ({pbs} ∩ fv iM) in nf(iN) ]]
                                                  && \text{since $[[ [mu] nf(iM) ]] = [[nf(iN)]]$ is proved } \\
                                                  &= [[ ord ({pas} ∩ fv iN) in nf(iN) ]]
                                                  && \text{because $\mu$ is a bijection between
                                                     $[[{pas} ∩ fv iN]]$ and $[[{pbs} ∩ fv iM]]$} \\
                                                  &= [[ ord {pas} in nf(iN) ]]
                                                  && \text{from \cref{lemma:fv-nf,corollary:ord-weakening} } \\
                                                  &= [[ pas' ]]
                                                  && \text{by the definition of $[[pas']]$} \\
      \end{aligned}$
    \end{enumerate}
  \item {\nameref{\ottdruleEOneExistsLabel}} Same as for \cref{case:ord-completeness:forall}.
  \item Other rules are congruent, and thus, proved by the corresponding congruent alpha-equivalence rule,
    which is applicable by the induction hypothesis. 
  \end{caseof}
\end{proof}

\begin{lemma}[Idempotence of normalization]
  \label{lemma:norm-idemp}
  Normalization is idempotent
  \begin{itemize}
  \item[$-$] $[[nf(nf(iN))]] = [[nf(iN)]]$
  \item[$+$] $[[nf(nf(iP))]] = [[nf(iP)]]$
  \end{itemize}
\end{lemma}
\begin{proof}
  By applying \cref{lemma:normalization-completeness} to \cref{lemma:normalization-soundness}.
\end{proof}

\begin{lemma}
  \label{lemma:normal-after-subst}
  The result of a substitution is normalized if and only if the initial type and
  the substitution are normalized.

  Suppose that $[[σ]]$ is a substitution  $[[Γ2 ⊢ σ : Γ1]]$,
  $[[iP]]$ is a positive type ($[[Γ1 ⊢ iP]]$),
  $[[iN]]$ is a negative type ($[[Γ1 ⊢ iN]]$). Then
  \begin{itemize}
  \item[$+$]
    $[[ [σ]iP  ]] \text{ is normal} \iff
    \begin{cases}
      [[σ|fv(iP)]] &\text{is normal}\\
      [[iP]]       &\text{is normal}\\
    \end{cases} $
  \item[$-$]
    $[[ [σ]iN  ]] \text{is normal} \iff
    \begin{cases}
      [[σ|fv(iN)]] &\text{is normal}\\
      [[iN]]       &\text{is normal}\\
    \end{cases} $
  \end{itemize}
\end{lemma}
\begin{proof}
  Mutual induction on $[[Γ1 ⊢ iP]]$ and $[[Γ1 ⊢ iN]]$.
  \begin{caseof}
  \item $[[iN]] = [[na]]$\\
    Then $[[iN]]$ is always normal, and
    the normality of $[[σ|{na}]]$ by the definition means $[[ [σ]na ]]$ is normal.

  \item $[[iN]] = [[iP → iM]]$\\
    \label{case:normal-after-subst-arrow}
    $
    \begin{aligned}[t]
      [[ [σ](iP → iM) ]] \text{ is normal} &\iff [[ [σ]iP → [σ]iM ]] \text{ is normal}
                                           && \text{by the substitution
                                              congruence} \\
                                           &\iff
                                             \begin{cases}
                                             [[ [σ]iP ]] &\text{is normal} \\
                                             [[ [σ]iM ]] &\text{is normal} \\
                                             \end{cases}\\
                                           &\iff
                                             \begin{cases}
                                               [[ iP ]]       &\text{is normal} \\
                                               [[ σ|fv(iP) ]] &\text{is normal} \\
                                               [[ iM ]]       &\text{is normal} \\
                                               [[ σ|fv(iM) ]] &\text{is normal} \\
                                             \end{cases}
                                           && \text{by the induction hypothesis}\\
                                           &\iff
                                             \begin{cases}
                                               [[ iP → iM ]]  &\text{is normal} \\
                                               [[ σ|fv(iP) ∪ fv(iM)]] &\text{is normal} \\
                                             \end{cases}\\
                                           &\iff
                                             \begin{cases}
                                               [[ iP → iM ]]  &\text{is normal} \\
                                               [[ σ|fv(iP→iM)]] &\text{is normal} \\
                                             \end{cases}
    \end{aligned}
    $
  \item $[[iN]] = [[↑iP]]$\\
    By congruence and the inductive hypothesis, similar to \cref{case:normal-after-subst-arrow}
  \item $[[iN]] = [[∀pas.iM]]$\\
    $
    \begin{aligned}[t]
      [[ [σ](∀pas.iM) ]] \text{ is normal} &\iff [[ (∀pas.[σ]iM) ]] \text{ is normal}
                                           && \text{assuming $[[pas]] \cap [[Γ1]] = \emptyset$ and
                                              $[[pas]] \cap [[Γ2]] = \emptyset$} \\
                                           &\iff
                                             \begin{cases}
                                             [[ [σ]iM ]] \text{ is normal} \\
                                             [[ord {pas} in [σ]iM = pas]] \\
                                             \end{cases}
                                           && \text{by the definition of normalization}\\
                                           &\iff
                                             \begin{cases}
                                               [[ [σ]iM ]] \text{ is normal} \\
                                               [[ord {pas} in iM = pas]] \\
                                             \end{cases}
                                           && \text{by \cref{lemma:ord-sigma}}\\
                                           &\iff
                                             \begin{cases}
                                               [[ σ|fv(iM) ]] \text{ is normal} \\
                                               [[ iM ]] \text{ is normal} \\
                                               [[ord {pas} in iM = pas]] \\
                                             \end{cases}
                                           && \text{by the induction hypothesis}\\
                                           &\iff
                                             \begin{cases}
                                               [[ σ|fv(∀pas.iM) ]] \text{ is normal} \\
                                               [[ ∀pas.iM ]] \text{ is normal} \\
                                             \end{cases}
                                           &&
                                              \begin{aligned}
                                              &\text{since $[[fv(∀pas.iM) = fv(iM)]]$;}\\ &\text{by the definition of normalization}
                                              \end{aligned}
    \end{aligned}
    $
  \item $[[iP]] = \dots$\\
    The positive cases are done in the same way as the negative ones.

  \end{caseof}
\end{proof}

>>>>

\begin{lemma}
  \label{lemma:uv-nf}
  Algorithmic variables are not changed by the normalization
  \begin{itemize}
  \item[$-$] $[[uv uN]] \equiv [[uv nf(uN)]]$
  \item[$+$] $[[uv uP]] \equiv [[uv nf(uP)]]$
  \end{itemize}
\end{lemma}
\begin{proof}
  By straightforward induction on $[[uN]]$ and mutually on $[[uP]]$, 
  similar to the proof of \cref{lemma:fv-nf}.
\end{proof}

@@

### 2023-09-05 proofs/tex/norm-lemmas.tex
Rearrange proofs

\begin{lemma}
  \label{lemma:uv-nf}
  Algorithmic variables are not changed by the normalization
  \begin{itemize}
  \item[$-$] $[[uv uN]] \equiv [[uv nf(uN)]]$
  \item[$+$] $[[uv uP]] \equiv [[uv nf(uP)]]$
  \end{itemize}
\end{lemma}
\begin{proof}
  By straightforward induction on $[[uN]]$ and mutually on $[[uP]]$.
\end{proof}

@@

\begin{lemma}[Soundness of normalization of algorithmic types]
  \label{lemma:normalization-soundness-alg}
  \hfill
  \begin{itemize}
    \item[$-$] $[[uN ≈ nf(uN)]]$
    \item[$+$] $[[uP ≈ nf(uP)]]$
  \end{itemize}
\end{lemma}
\begin{proof}
  The proof coincides with the proof of \cref{lemma:normalization-soundness}.
\end{proof}

\begin{corollary}[Normalization preserves ordering]
  \label{corollary:normalization-ord}
  For any $[[varset]]$,
  \begin{itemize}
  \item[$-$] $[[ord varset in nf(uN)]] = [[ord varset in uM]]$
  \item[$+$] $[[ord varset in nf(uP)]] = [[ord varset in uQ]]$
  \end{itemize}
\end{corollary}
\begin{proof}
  Immediately from \cref{lemma:ord-completeness,lemma:normalization-soundness}.
\end{proof}

@@

  \begin{itemize}
  \item[$-$] If $[[iN ≈ iM]]$ then $[[nf(iN)]] = [[nf(iM)]]$
  \item[$+$] If $[[iP ≈ iQ]]$ then $[[nf(iP)]] = [[nf(iQ)]]$
  \end{itemize}
  (Here equality means alpha-equivalence)
\end{lemma}

>>>>

@@

\begin{corollary}[Normalization preserves ordering]
  \label{corollary:normalization-ord}
  For any $[[varset]]$,
  \begin{itemize}
  \item[$-$] $[[ord varset in nf(iN)]] = [[ord varset in iM]]$
  \item[$+$] $[[ord varset in nf(iP)]] = [[ord varset in iQ]]$
  \end{itemize}
\end{corollary}
\begin{proof}
  Immediately from \cref{lemma:ord-completeness,lemma:normalization-soundness}.
\end{proof}

@@

  \begin{itemize}
  \item[$-$] If $[[iN ≈ iM]]$ then $[[nf(iN)]] = [[nf(iM)]]$
  \item[$+$] If $[[iP ≈ iQ]]$ then $[[nf(iP)]] = [[nf(iQ)]]$
  \end{itemize}
\end{lemma}

### 2023-09-05 proofs/tex/ord-lemmas.tex
Rearrange proofs

\begin{lemma}[Soundness of variable ordering]
  \label{lemma:ord-soundness}
  Variable ordering extracts precisely used free variables.
  \begin{itemize}
    \item[$-$] $[[ {ord varset in uN} ]] \equiv [[varset ∩ fv uN]]$ (as sets)
    \item[$+$] $[[ {ord varset in uP} ]] \equiv [[varset ∩ fv uP]]$ (as sets)
  \end{itemize}
\end{lemma}
\begin{proof}
  Straightforward mutual induction on 
  $[[ ord varset in uN = ordVars ]]$ and $[[ ord varset in uP = ordVars ]]$
\end{proof}

@@

\begin{lemma}[Completeness of variable ordering]
  \label{lemma:ord-completeness}
  Variable ordering is invariant under equivalence. For arbitrary $[[varset]]$,
   \begin{itemize}
  \item[$-$] If $[[iN ≈ iM]]$ then $[[ord varset in uN]] = [[ord varset in uM]]$ (as lists)
  \item[$+$] If $[[iP ≈ iQ]]$ then $[[ord varset in uP]] = [[ord varset in uQ]]$ (as lists)
  \end{itemize}
\end{lemma}
\begin{proof}
  Mutual induction on $[[iN ≈ iM]]$ and $[[iP ≈ iQ]]$.
\end{proof}

>>>>

\begin{lemma}[Soundness of variable ordering]
  \label{lemma:ord-soundness}
  Variable ordering extracts used free variables.
  \begin{itemize}
    \item[$-$] $[[ {ord varset in iN} ]] = [[varset ∩ fv iN]]$ (as sets)
    \item[$+$] $[[ {ord varset in iP} ]] = [[varset ∩ fv iP]]$ (as sets)
  \end{itemize}
\end{lemma}
\begin{proof}
  Straightforward mutual induction on 
  $[[ ord varset in iN = ordVars ]]$ and $[[ ord varset in iP = ordVars ]]$
\end{proof}

@@

\begin{lemma}[Completeness of variable ordering]
  \label{lemma:ord-completeness}
  Variable ordering is invariant under equivalence. For arbitrary $[[varset]]$,
   \begin{itemize}
  \item[$-$] If $[[iN ≈ iM]]$ then $[[ord varset in iN]] = [[ord varset in iM]]$ (as lists)
  \item[$+$] If $[[iP ≈ iQ]]$ then $[[ord varset in iP]] = [[ord varset in iQ]]$ (as lists)
  \end{itemize}
\end{lemma}
\begin{proof}
  Mutual induction on $[[iN ≈ iM]]$ and $[[iP ≈ iQ]]$.
\end{proof}
